---
title: "Project-01"
output: pdf_document
author: jose Antonio (joseantonio@me.com)
---

github: [Project](https://github.com/joseantonio11/coursera-statistical-inference)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require('knitr', quietly=TRUE)
```

## I- TITLE: Simulating Exponential Distribution

## II - OVERVIEW:

In this project we use the exponential distribution in R and compare it with the Central Limit Theorem. The following criteria will be used for this project:

- mean of exponential distribution = 1/lambda
- standard deviation = 1/lambda
- lambda = 0.2 for simulations
- investigate distribution of 40 exponentials (will need to do 1000 simulations)

The problems to se solved are:

1 - Show the sample mean and compare it to the theoretical mean of the distribution.

2 - Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

3 - Show that the distribution is approximately normal.

### Libraries
```{r, echo=TRUE}
getwd()
library(knitr)
library(markdown)
library(rmarkdown)
library(plyr)
library(stats)
library(ggplot2)
```

## III - SIMULATIONS
Answering the questions using R and explanations

## 1 - Sample Mean versus Theoretical Mean

```{r, echo=TRUE}
set.seed(1) ## reproducible results
lambda <- 0.2 # rate for simulations
n <- 40 # number of samples
simulations <- 1:1000 # suggested
meansd <- 1/0.2 # mean of exponential distribution
sdexp <- 1/0.2 # standard deviatin of exponential distribution
```

### 1.1 - Sample Mean

```{r,echo=TRUE}
sample_mean <- data.frame(simcol=sapply(simulations, function(simcol) 
        {mean(rexp(n,lambda))}))
```

### 1.2 - Center Sample Distribution

```{r, echo=TRUE}
mean(sample_mean$simcol)
```

### 1.3 - Center Theoretical Distribution

```{r, echo=TRUE}
1/lambda
```

* THE SAMPLE MEAN (4,99) IS VERY CLOSE TO THEORETICAL MEAN (5)

### 1.4 - Histogram

```{r, echo=TRUE}
hist(sample_mean$simcol, main = "1000 averages of 40 random", cex.main=0.7)
abline(v=mean(sample_mean$simcol),col="green")
abline(v=1/lambda,col="red")
```


## 2 - Sample Variance versus Theoretical Variance

### 2.1 - Calculating the standard deviation (sample and theorical distribution)

```{r, echo=TRU}
## sd of sample distribution
sd(sample_mean$simcol)
## sd of theoretical distribution
(1/lambda)/sqrt(40)
```

## 2.2 - Calculating Variance (sample and theorical distribution)

```{r, echo=TRUE}
## Variance of sample distribution
var(sample_mean$simcol)
## Variance of theoretical distribution
((1/lambda)/sqrt(40))^2
```

* THE SAMPLE SD IS CLOSE TO THEORETICAL SD. THE VARIANCE IS SIMILAR

## 2.3 - Histogram

```{r, echo=TRUE}
hist(sample_mean$simcol, main = "1000 averages of 40 random", xlab = "Mean", cex.main=0.7)
abline(v=mean(sample_mean$simcol),col="blue")
## Standard Deviation
abline(v=mean(sample_mean$simcol)+ sd(sample_mean$simcol), col="red", lw=2)
abline(v=mean(sample_mean$simcol)- sd(sample_mean$simcol), col="red", lw=2)
## Theoretical SD
abline(v=mean(sample_mean$simcol)+sqrt(sdexp^2/n), col = "green")
abline(v=mean(sample_mean$simcol)-sqrt(sdexp^2/n), col = "green")
```

## 3 - Distribution







```{r, echo=TRUE}
cat("Sample mean = ", s-mean, "  ", "Theoretical Distribution= ", center-dist)
```



